The role of advection versus diffusion in intraventricular transport

### 2025-03-06 paper/main.tex
Updates with MER and HH

\subsection{Absence of ciliary motion affects local protein distribution}
Since CSF flow is highly disturbed in the absence of motile cilia, we
want to assess the impact of cilia motion on solute distribution. To this end,
we measured the movement of photoconverted Dendra2 signal
in control and motile-cilia mutant (\emph{smh}) zebrafish .
By plotting the means of $\Delta F$ (\Crefpanels{fig:fig4}{a})
and the time to threshold
across the control and mutant cohorts (\Crefpanels{fig:fig4}{b}),
we identified a significant
difference in Dendra2 distribution in ROIs 2 and 4, but not in ROIs 3, 5, and 6. In ROIs 2 and 4,
which measure Dendra2 distribution downstream of the photoconversion site,
we observed a slower increase of fluorescence intensity,
resulting in $45\%$ (ROI 2) and $25\%$ (ROI 4)
higher mean times-to-threshold in the \emph{smh}
mutants compared to controls (\Crefpanels{fig:fig4}{b}).

Next, to identify whether our model would recapitulate these findings,
we simulate protein transport by using the only-cardiac/no-cilia flow model 
and compare it with the baseline model
(both cilia and cardiac motion) (\Crefpanels{fig:fig4}{c}).
Similar to our experiments, we observe slower dynamics in simulations for ROIs 2 and 4,
but small changes for ROIs 3 and 6.
Interestingly, the difference was even greater for the simulations,
with the times-to-threshold in ROI 2 and 4 being 90\% and 34\% higher
in the absence of cilia-mediated flow.
Moreover, ROI 5 exhibits very different $\cbar$ profiles for the two
configurations in our computational model,
but not in our experimental data.
Altogether, our findings identify that diffusion plays an important
role for solute movement to regions upstream in the cilia flow fields or to distant ventricles.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F(t)$
    in each region of interest (ROI) for control (black lines)
    and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets. The $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%, with 
    the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}

>>>>

\subsection{Absence of ciliary motion affects local protein distribution}

\mer{Sentence on why absence of motile cilia is intersting: Cilia
  motion is key to development and absence of cilia is associated with
  pathologies.} As observed both experimentally and computationally,
the absence of motile cilia alters the CSF flow patterns. We next
investigate the implications of these changes on intraventricular
transport.

We first experimentally measured the movement of photoconverted
Dendra2 signal in control and motile-cilia mutant (\emph{smh})
zebrafish. In ROIs 2 and 4, which measure Dendra2 distribution
downstream of the photoconversion site, we observed a slower increase
of fluorescence intensity (\Crefpanels{fig:fig4}{a}), resulting in
$45\%$ (ROI 2) and $25\%$ (ROI 4) higher mean times-to-threshold
(\Crefpanels{fig:fig4}{b}), in the \emph{smh} mutants compared to
controls. We found no significant differences in the Dendra2
distribution in ROIs 3, 5, and 6 between \emph{smh} mutants and
controls.

We next mimic the \emph{smh} mutants in the computational model by
considering transport of Dendra2 with CSF flow driven by cardiac
pressure pulsations only, representing the absence of motile
cilia. Comparing the in-silico mutants with transport predictions in
the presence of both cardiac pulsations and motile cilia, we observe
noticeably delayed dynamics in ROIs 2, 4 and 5, but little change for
ROIs 3 and 6 (\Crefpanels{fig:fig4}{c}).

\mer{MER: MER and HH stopped here.}

Comparing with the experiment, the difference between the control and
mutant scenarios was even greater for the simulations, with the
times-to-threshold in ROI 2 and 4 being 90\% and 34\% higher in the
absence of cilia-mediated flow. Moreover, ROI 5 exhibits very
different $\cbar$ profiles for the two configurations in our
computational model, but not in our experimental data.  Altogether,
our findings identify that diffusion plays an important role for
solute movement to regions upstream in the cilia flow fields or to
distant ventricles.

%% Again, the simulated Dendra2 distribution patterns differ from the

%% Next, to identify whether our model would recapitulate these findings,
%% we simulate CSF flow driven only by cardiac pressure pulsations and
%% the resulting diffusion and advection of

%% Dendra2 transport

%% by using the only-cardiac/no-cilia flow model 
%% and compare it with the baseline model
%% (both cilia and cardiac motion) (\Crefpanels{fig:fig4}{c}).

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure4_compare_control_mutant.png}
    \caption{\textbf{a.} Relative change in fluorescence intensity $F(t)$
    in each region of interest (ROI) for control (black lines)
    and \emph{smh} mutant (red lines) fish. 
    \textbf{b.} Times when $\Delta F(t)$ first exceeds a threshold value,
    equal to 0.25 (ROIs 2--4) or 0.10 (ROIs 5, 6), for control (gray dots) and
    \emph{smh} mutant (red dots) fish.
    The thick black lines denote mean values of the datasets. The $p$-values
    were calculated using a Mann-Whitney U test with a confidence level of 95\%, with 
    the null hypothesis that the two distributions of control and mutant data are 
    from the same population.
    \textbf{c.} Mean concentration $\cbar(t)$ in each ROI,
    simulated using the baseline model (black lines) and the cardiac-only/no-cilia model (red lines).
    The horizontal gray lines mark the threshold value $\hat{c}$, and the time
    $\hat{t}$ when the threshold is first exceeded is reported in the plot legends.
    Note that in both panels \textbf{a} and \textbf{c}, the vertical axes have
    different limits in each plot.}
    \label{fig:fig4}
\end{figure}
